### 2025-01-26 analysis/preprocessing.qmd
chore(preprocessing): update classification output
-aggregate model performance across valence and arousal for bhuvanakumar2023em &  nguyen20172017an -update model categories for zhang2017fe

```{r}
#| eval: true
#| echo: false
# Classify techniques according Hastie, Tibshirani, Friedman (2008)
# https://www.sas.upenn.edu/~fdiebold/NoHesitations/BookAdvanced.pdf
#
metaMER_results$model_id <- tolower(metaMER_results$model_id)
metaMER_results$model_class_id <- 'Unclassified'
metaMER_results$model_class_id[str_detect(metaMER_results$model_id,'lr|lm|pls|mlr|pcr|logistic regression|2d model full|pentagon|gaussian process regression|sparse bayesian regression|variational bayesian regression|logistic regression|lda|rda|regularized discriminant analysis|reguliarized discriminant analysis')] <- 'Linear Methods' # Class name from Elements of Stat.."
#metaMER_results$model_class_id[str_detect(metaMER_results$model_id,'logistic regression|lda|rda|regularized discriminant analysis|reguliarized discriminant analysis')] <- 'Linear Classification'
#metaMER_results$model_class_id[str_detect(metaMER_results$model_id,'mars|gam')]<-'Additive Trees and Related Methods' #
metaMER_results$model_class_id[str_detect(metaMER_results$model_id,'rbf|gmm|local|polynomial|polygonal|knn|mars|gam')]<-'Kernel Smoothing, Additive and KNN'
#metaMER_results$model_class_id[str_detect(metaMER_results$model_id,'adaboost|gradient')]<-'Boosting'
metaMER_results$model_class_id[str_detect(metaMER_results$model_id,'nn|gru|lstm|ltsm|long short term memory|rprop|mcan')]<-'Neural Nets'
metaMER_results$model_class_id[str_detect(metaMER_results$model_id,'svm|svr|support vector regression|smoreg|smo ')]<-'Flexible Discriminants'
#metaMER_results$model_class_id[str_detect(metaMER_results$model_id,'knn')]<-'Prototype methods'
metaMER_results$model_class_id[str_detect(metaMER_results$model_id,'rf|extremely randomized tree regression|random forest|adaboost|gradient')]<-'Random Forests'
#metaMER_results$model_class_id[str_detect(metaMER_results$model_id,'rf')]<-'Ensemble Learning'

@@

cat(paste("\nWhere", length(unique(metaMER_results$citekey[metaMER_results$model_category=='classification'])), "are classification studies"))
# note that we have some classification studies that also do regression and vice versa?
# THIS IS CORRECT (updated 24 October 2024):
# THIS IS CORRECT (updated 17 January 2025):
# [1] "We have 1114 observations"
# [1] "We have 34 studies"
# [1] "Where 22 are regression studies"
# [1] "Where 12 are classification studies"

@@

@@

C_studies$citekey <- factor(C_studies$citekey)

C_summary <- summarise(group_by(C_studies,citekey),valuesMean=mean(values,na.rm=TRUE),valuesMedian=median(values,na.rm=TRUE),valuesMax=max(values,na.rm=TRUE),stimulus_n=first(stimulus_n),studyREF=first(studyREF),model_class_id=first(model_class_id), stimulus_genre_mixed=first(stimulus_genre_mixed),journal_type = first(journal_type))

@@

@@

tmp <- drop_na(C_studies)
tmp$dimension<-str_to_title(tmp$dimension)
tmp$model_class_id<-factor(tmp$model_class_id,
                           levels = c("Neural Nets","Flexible Discriminants", "Kernel Smoothing, Additive and KNN", "Random Forests","Linear Methods"),
                           labels = c("Neural\nNets","Flexible\nDiscriminants", "KS\n & KNN", "Random\nForests", "Linear\nMethods"))

>>>>

```{r}
#| eval: true
#| echo: false
# Classify techniques according Hastie, Tibshirani, Friedman (2008)
# https://www.sas.upenn.edu/~fdiebold/NoHesitations/BookAdvanced.pdf
#
metaMER_results$model_id <- tolower(metaMER_results$model_id)
metaMER_results$model_class_id <- 'Unclassified'
metaMER_results$model_class_id[str_detect(metaMER_results$model_id,'lr|lm|pls|mlr|pcr|logistic regression|2d model full|pentagon|gaussian process regression|sparse bayesian regression|variational bayesian regression|logistic regression|lda|rda|regularized discriminant analysis|reguliarized discriminant analysis|linear discriminant|quadratic discriminant')] <- 'Linear Methods' # Class name from Elements of Stat.."
#metaMER_results$model_class_id[str_detect(metaMER_results$model_id,'logistic regression|lda|rda|regularized discriminant analysis|reguliarized discriminant analysis')] <- 'Linear Classification'
#metaMER_results$model_class_id[str_detect(metaMER_results$model_id,'mars|gam')]<-'Additive Trees and Related Methods' #
metaMER_results$model_class_id[str_detect(metaMER_results$model_id,'rbf|gmm|local|polynomial|polygonal|knn|mars|gam')]<-'Kernel Smoothing, Additive and KNN'
#metaMER_results$model_class_id[str_detect(metaMER_results$model_id,'adaboost|gradient')]<-'Boosting'
metaMER_results$model_class_id[str_detect(metaMER_results$model_id,'nn|gru|lstm|ltsm|long short term memory|rprop|mcan')]<-'Neural Nets'
metaMER_results$model_class_id[str_detect(metaMER_results$model_id,'svm|svr|support vector regression|smoreg|smo ')]<-'Flexible Discriminants'
#metaMER_results$model_class_id[str_detect(metaMER_results$model_id,'knn')]<-'Prototype methods'
metaMER_results$model_class_id[str_detect(metaMER_results$model_id,'rf|extremely randomized tree regression|random forest|adaboost|gradient|tree|bag|boost')]<-'Random Forests'
#metaMER_results$model_class_id[str_detect(metaMER_results$model_id,'rf')]<-'Ensemble Learning'

@@

cat(paste("\nWhere", length(unique(metaMER_results$citekey[metaMER_results$model_category=='classification'])), "are classification studies"))
# note that we have some classification studies that also do regression and vice versa?
# THIS IS CORRECT (updated 17 January 2025):
# [1] "We have 1114 observations"
# [1] "We have 34 studies"
# [1] "Where 22 are regression studies"
# [1] "Where 12 are classification studies"

@@

# a combination of the two
R_studies$feature_n_complexity_genre <- cut(R_studies$feature_n,
                                      breaks = c(0,15,501,653,20000),
                                      labels = c("Small single genre study","Medium single genre/multigenre study","Medium-large single genre/multigenre study","Huge multigenre study"))
#table(R_studies$feature_n_complexity_genre)
```

@@

@@

```{r}
#| echo: false
#| eval: false
C_studies$citekey <- factor(C_studies$citekey)
tmp <- drop_na(C_studies)
tmp$dimension<-str_to_title(tmp$dimension)
tmp$model_class_id<-factor(tmp$model_class_id,
                           levels = c("Neural Nets","Flexible Discriminants", "Kernel Smoothing, Additive and KNN", "Random Forests","Linear Methods"),
                           labels = c("Neural\nNets","Flexible\nDiscriminants", "KS\n & KNN", "Random\nForests", "Linear\nMethods"))

## Aggregate valence and arousal results for nguygen2017, bhuvanakumar2023

```{r}
C_summary <- summarise(group_by(C_studies,dimension,citekey),valuesMean=mean(values,na.rm=TRUE),valuesMedian=median(values,na.rm=TRUE),valuesMax=max(values,na.rm=TRUE),stimulus_n=first(stimulus_n),studyREF=first(studyREF),model_class_id=first(model_class_id),feature_n=first(feature_n),journal_type=first(journal_type),feature_n=first(feature_n),feature_n_complexity=first(feature_n_complexity),stimulus_genre_mixed=first(stimulus_genre_mixed)#,
                        #feature_n_complexity_genre=first(feature_n_complexity_genre)
                        )
# 2025-01-24: averaging valence and arousal for this study, which makes the max equivalent to both the mean and median (two models: one valence, one arousal)
C_summary[C_summary$citekey == "nguyen2017an",]$valuesMax <- C_summary[C_summary$citekey == "nguyen2017an",]$valuesMean

# 2025-01-24: averaging across quadrants for bhuvanakumar study
# subset study
bhuvanakumar <- subset(
  C_studies, 
  stringr::str_detect(
    C_studies$citekey,
    "bhuvanakumar"
    )
)
# group by algorithm to average across the four quadrants
bhuvanakumar$model_algorithm <- substr(
  bhuvanakumar$model_id, 
  1, 
  4
)
# take the average for each algorithm
bhuvanakumar |> 
  group_by(model_algorithm) |>
  mutate(values = mean(values)) -> bhuvanakumar
# overwrite values in C_summary
C_summary[C_summary$citekey == "bhuvanakumar2023em",]$valuesMean <- mean(unique(bhuvanakumar$values))
C_summary[C_summary$citekey == "bhuvanakumar2023em",]$valuesMedian <- median(unique(bhuvanakumar$values))
C_summary[C_summary$citekey == "bhuvanakumar2023em",]$valuesMax <- max(unique(bhuvanakumar$values))

remove(bhuvanakumar)

@@
